---
title: "example_data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{nmrrr_data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nmrrr)
```

This package contains two example datasets:
1. kfp_hysteresis
1. kfp_drydown

Additional datasets can be found in the GitHub repository for the nmrrr package:
1. amp_tempest
1. meb_burn



These are much smaller subsets of larger datasets generated as part of larger experiments.


## `kfp_hysteresis` dataset

This dataset contains ten samples from [Patel et al. 2021](https://doi.org/10.1016/j.soilbio.2021.108165).
Sample 29 represents a drought treatment, and sample 34 represents a flood treatment.  

**Sample preparation:**
Soluble organic carbon was extracted from soil samples using deionized MilliQ water; 
these extracts were then freeze-dried and reconstituted in DMSO-D6 for 1H-NMR analysis. 

**Solvent:**
DMSO-D6

**Spectra cleaned using:**
MNova


## `kfp_drydown` dataset

**Sample preparation:**
Soluble organic carbon was extracted from soil samples using deionized MilliQ water; 
these extracts were then freeze-dried and reconstituted in D2O for 1H-NMR analysis. 

**Solvent:**
D2O

**Spectra cleaned using:**
MNova


## `amp_tempest` dataset

**Sample preparation:**

**Solvent:**
MeOD

**Spectra cleaned using:**
MNova  
Note: Peaks are exported in single-column format


## `meb_burn` dataset

**Sample preparation:**

**Solvent:**
MeOD

**Spectra cleaned using:**
TopSpin  
