---
title: "Bin sets included with nmrrr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bin sets included with nmrrr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nmrrr)
```

The `{nmrrr}` package provides some bin sets for compound classification.

### `bins_Clemente2012`

NMR grouping bins from Clemente et al. (2012), using DMSO-D6 as solvent. 

1. aliphatic polymethylene and methyl groups (0.6–1.3 ppm, “aliphatic1”); 
2. aliphatic methyl and methylene near O and N (1.3–2.9 ppm, “aliphatic2”); 
3. O-alkyl, mainly from carbohydrates and lignin (2.9–4.1 ppm); 
4. α-proton of peptides (4.1–4.8 ppm); 
5. aromatic and phenolic (6.2–7.8 ppm); and 
6. amide, from proteins (7.8–8.4 ppm).

**Source**

JS Clemente et al. 2012. “Comparison of Nuclear Magnetic Resonance Methods for the Analysis of Organic Matter Composition from Soil Density and Particle Fractions.” Environmental Chemistry. https://doi.org/10.1071/EN11096.



### `bins_Hertkorn2013`

NMR grouping bins from Hertkorn et al. (2013), using MeOD as solvent. 

1. aliphatics, HCCC (0.0-1.9); 
2. acetate analogs and CRAM (carboxyl-rich alicyclic materials), HCX (1.9-3.1); 
3. carbohydrate-like and methoxy, HCO (3.1-4.9); 
4. olefinic HC=C (5.3-7.0); 
5. aromatic (7.0-10.0).

**Source**

N. Hertkorn et al. 2013. "High-field NMR spectroscopy and FTICR mass spectrometry: powerful discovery tools for the molecular level characterization of marine dissolved organic matter" Biogeosciences https://doi.org/10.5194/bg-10-1583-2013.


### `bins_Lynch2019`

NMR grouping bins from Lynch et al. (2019), using D2O as solvent. 

1. methyl, methylene, and methane bearing protons (0.6–1.6 ppm) ; 
2. unsaturated functional groups (1.6–3.2 ppm), including ketone, benzylic, and alicyclic-bearing protons; 
3. unsaturated, heteroatomic compounds, including O-bearing carbohydrates, ethers, and alcohols (3.2–4.5 ppm); 
4. conjugated, double bond functionalities, including aromatic, amide, and phenolic structures (6.5–8.5 ppm).

**Source**

LM Lynch et al. 2019. “Dissolved Organic Matter Chemistry and Transport along an Arctic Tundra Hillslope.” Global Biogeochemical Cycles https://doi.org/10.1029/2018GB006030.

### `bins_Mitchell2018`

NMR grouping bins from Mitchell et al. (2018), using DMSO-D6 as solvent. 

1. aliphatic polymethylene and methyl groups (0.6–1.3 ppm); 
2. N- and O-substituted aliphatic (1.3–2.9 ppm); 
3. O-alkyl (2.9–4.1 ppm); 
4. α-proton of peptides (4.1–4.8 ppm); 
5. anomeric proton of carbohydrates (4.8–5.2 ppm); 
6. aromatic and phenolic (6.2–7.8 ppm); 
7. amide (7.8–8.4 ppm).

**Source**

P Mitchell et al. 2018. “Nuclear Magnetic Resonance Analysis of Changes in Dissolved Organic Matter Composition with Successive Layering on Clay Mineral Surfaces.” Soil Systems https://doi.org/10.3390/soils2010008.


## Using your own bin sets

[TEXT COMING SOON]
